---
title: "Week 03"
subtitle: "Data wrangling pt. 1"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
```


### Announcements

- Notecard due at end of class. Answer one (or more):
    + What was most confusing/unclear?
    + What questions are you wondering about?
    + How can I be praying for you? 
- AE 03 (in-class today) due Thursday by class time (4:20pm)
- First 2 [Statistics Experience points](https://kgfitzgerald.github.io/APU_MATH_250/hw/se-01.html) due next Thursday, September 22
- HW 01 due next Thursday, September 22
    + Encouraged to attempt all exercises on your own, before seeking external help
    + Will help you gauge your current level of understanding and hone your trouble-shooting skills ahead of Exam 1

### Questions?

*What's the difference between the `fill` and `color` aesthetics in `ggplot`?*

+ For histograms / things with bars, `color` = outside, `fill` = inside.
```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(bins = 10, color = "blue")

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(bins = 10, color = "blue", fill = "red")
```

+ For points, use `color`
```{r}
ggplot(midwest, aes(x = area, y = poptotal, 
                    color = state)) +
  geom_point()
```

Note how the above examples set `color` inside or outside `aes()` in the global `ggplot()` layer.

+ Put inside `aes()` if you want to map a variable in the data to the color/fill
+ Put outside `aes()`, usually in the `geom_` layer, if you want to set `color/fill` to a specific color for the whole plot
+ Later layers will override earlier layers

```{r}
ggplot(midwest, aes(x = area, y = poptotal, 
                    color = state)) +
  geom_point(color = "pink")
```

*What does data wrangling mean? It seems from the videos that we are employing certain commands in r in order to focus on specific aspects of data, but is that the essence of "wrangling" data? Or does it just mean making it more manageable for analysis?*

+ Yes to both :) wrangle = manipulate data into a form so that your question can be answered / task can be accomplished

*How do I suppress warnings and messages*?

+ Add `message = FALSE, warning = FALSE` inside the curly brackets of a code chunk. 

### Application Exercise

- The remainder of class will be spent on AE-03. 
- You can access it from your RStudio Cloud account. 
- It is due by class-time on Thursday. 
- To turn it in, you should upload your .html file to Canvas.