---
title: "Week 09"
subtitle: "Logistic Regression"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Thursday Announcements

-   Due tonight:

    -   AE 07

    -   Lab 07

    -   Project data cleaning & EDA

-   Today:

    -   Data viz mini-workshop

    -   Project work session

-   Next week:

    -   Lecture videos, but no prep quiz

    -   HW 04 due Thursday

### Tuesday Announcements

-   Today:
    -   Finish Lab 07 (w/ interaction model), due Thursday
    -   AE-07 (due Thursday, midnight)
-   Thursday:
    -   Data viz best practices mini-workshop
    -   Project work session
    -   Project data cleaning & EDA due Thursday

### Questions?

-   I generally understand the idea of main effects vs interaction-effects model, but how do we determine which one is a better model to use?
-   I still don't quite understand when we should and shouldn't use `factor()` in plots

```{r, message = FALSE, warning = FALSE}
library(openintro)
library(tidyverse)

evals <- evals %>% 
  mutate(gender_numeric = as.numeric(gender))

ggplot(evals, aes(x = bty_avg, y = score, 
                  color = gender_numeric)) +
  geom_point()

ggplot(evals, aes(x = bty_avg, y = score, 
                  color = factor(gender_numeric))) +
  geom_point()
```

-   Project timeline? Will we have time to work on projects in class?

+------------------+-------------------------------+----------------------+-----------------------------+
| Week/Dates       | Tue (in-class)                | Thur (in-class)      | Project Due                 |
+==================+===============================+======================+=============================+
| Week 09          | AE-07                         | Data vis workshop    | Data cleaning & EDA (Thurs) |
|                  |                               |                      |                             |
| (Oct 24 - 29)    |                               | Project work session |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| Week 10          | AE-08                         | Lab 08               |                             |
|                  |                               |                      |                             |
| (Oct 31 - Nov 5) |                               |                      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| Week 11          | AE-09                         | Lab 09               |                             |
|                  |                               |                      |                             |
| (Nov 7 - Nov 12) |                               |                      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| Week 12          | AE-10                         | Project work session | Rough draft (Thurs)         |
|                  |                               |                      |                             |
| (Nov 14 - 19)    |                               |                      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| THANKSGIVING     |                               |                      |                             |
|                  |                               |                      |                             |
| (Nov 21 - 26)    |                               |                      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| Week 13          | Special Topics: Shiny Apps    | Project Peer Review  |                             |
|                  |                               |                      |                             |
| (Nov 28 - Dec 3) | (AE-11)                       |                      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| Week 14          | Special Topics: Text analysis | Project Work Session |                             |
|                  |                               |                      |                             |
| (Dec 5 - 10)     | (AE-12)                       |                      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+
| Week 15          |                               | Presentations        |                             |
|                  |                               |                      |                             |
| (Dec 12 - 17)    |                               | (4:05 - 5:55pm)      |                             |
+------------------+-------------------------------+----------------------+-----------------------------+

-   Vote: prefer to switch Lab 09 and Project work session (in Weeks 11 and 12)?

-   Vote: special topics preference - Shiny apps or text analysis?

### Application Exercise

-   The remainder of class will be spent on AE-07.
-   You can access it from your RStudio Cloud account.
-   It is due by class-time on Thursday.
-   To turn it in, you should upload your .html file to Canvas.
