---
title: "Week 05"
subtitle: "Scientific Practice & Data Science Workflow"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
```


### Announcements


### Exam ground rules

+ This is an INDIVIDUAL EXAM. You may NOT work together or consult other people on this exam.
+ You are ENCOURAGED to use your notes, lecture slides/videos, previous assignments, and the internet
    + You may utilize already existing resources on the internet, but you may NOT post to any online forums to ask for help on the exam.  
    + If you find code online, cite it, and annotate the code to demonstrate understanding. 
+ Failure to adhere to any of the above policies will be considered a violation of academic integrity and will be handled according to university policy.

### Reminders / Tips

+ Make sure you've answered all parts of the question!
+ Make sure to clean up visualizations - add titles, clean axis labels, all text should be readable
+ Use `kbl()` for clean formatting of tables
+ Google your errors
+ Search "how to ____ in tidyverse r"
+ stackoverflow forum is usually particularly helpful
+ Check out textbooks linked in prep material (esp. R for Data Science)

### Questions?


### Project Work Session

- See [Doing Data Science slides](https://kgfitzgerald.github.io/APU_MATH_250/slides/16-doing-data-science.html#1) for getting started on your [project proposal](https://apumath250.netlify.app/project/#project-proposal)


