---
title: "Week 04"
subtitle: "Data wrangling II"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Scientist of the Week

[Mike Dairyko](https://hardin47.github.io/CURV/scholars/dairyko.html)

### Announcements


### Tidyverse style guidelines

[Tidyverse style guidelines](https://style.tidyverse.org/pipes.html) for line breaks, white space, etc.

1.  Have space before `%>%` and a linebreak after

```{r, eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarize_if(is.numeric, mean) %>%
  ungroup() %>%
  gather(measure, value, -Species) %>%
  arrange(value)

# Bad
iris %>% group_by(Species) %>% summarize_all(mean) %>%
ungroup %>% gather(measure, value, -Species) %>%
arrange(value)
```

2.  `ggplot()` layers each go on their own line, space before `+`, all but first layer indented.

```{r, eval = FALSE}
# Good
ggplot(aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() +
  labs(
    x = "Sepal width, in cm",
    y = "Sepal length, in cm",
    title = "Sepal length vs. width of irises"
  ) 

# Bad
ggplot(aes(x = Sepal.Width, y = Sepal.Length, color = Species))+
  geom_point() + labs(x = "Sepal width, in cm", y = "Sepal length, in cm", title = "Sepal length vs. width of irises") 
```

3.  If arguments to a function don't fit on one line, put each argument on its own line

```{r, eval = FALSE}
iris %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length),
    Sepal.Width = mean(Sepal.Width),
    Species = n_distinct(Species)
  )

iris %>%
  group_by(Species) %>%
  summarise(Sepal.Length = mean(Sepal.Length),
            Sepal.Width = mean(Sepal.Width),
            Species = n_distinct(Species))
```

4.  Spaces on each side of all operators (e.g. `+, =, &, <-,`) and after commas

```{r, eval = FALSE}
# Good
iris %>% 
  mutate(ratio = Sepal.Length/Sepal.Width) %>% 
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
  geom_point()

# Bad
iris%>% 
  mutate(ratio=Sepal.Length/Sepal.Width) %>% 
  ggplot(aes(x=Sepal.Width,y = Sepal.Length)) +
  geom_point()
```

### Questions

*Is there a way to debug code in RStudio? I'm not too familiar with it so whenever I get an error, I just go about the trial-and-error method until it works, but this can be a long and exhausting process.*

-   Google your errors :) stackoverflow.com typically has a lot of great help
-   use the `?` command with the name of the function to pull up the help documentation. E.g. `?mean` will pull up the documentation for the function `mean()`
-   Common things to look out for: spelling, capitalization, misplaced or missing commas/parentheses/quotations
-   Run all code chunks above (button next to play button) then run current code chunk
-   Check your Environment to make sure you have the objects you think you have, that you're spelling them correctly, and that they have the dimensions you're expecting
-   Run `View()` in your console to actually view/inspect your data. Can temporarily add this on to the end of a pipe so you can see your results more clearly.

### Reminder of Class Norms

-   Everyone has expertise
-   Be present
-   Participate to the best of your ability
-   Share talk time
-   Critique ideas, not people
-   Engage new perspectives with curiosity, not judgement
-   Exhibit intellectual humility
-   Embrace discomfort
-   Normalize time to think

#### What hinders group dynamics

-   Rely on others too much
-   Lack of preparation
-   Over/under participation
-   Lack of communication
-   Lack of confidence
-   Lack of patience
-   Lack of focus

#### What makes enriching group dynamics

-   Get to know classmates
-   Communication & Clarity
-   Problem solving with different approaches
-   Preparation, sleep
-   Accountability
-   Checking in
-   Defining roles / delegation
-   Clear instruction
-   Self-awareness

#### Recommendations

-   Don't leave class on Thursdays without discussing a plan for how/when Lab will be finished. Note: it should NOT fall on one person to do this.
-   Check in regularly that each group member is understanding each exercise / part of the code - don't move on until everyone is on the same page.
    -   Normalize question asking

### Application Exercise

-   The remainder of class will be spent on AE-04.
- You can access it from GitHub. 
- It is due at the end of class today. 
-   To turn it in, you should upload your .html file to Canvas.
