---
title: "Lab 07 - Grading the professor, Pt. 2"
subtitle: "Modelling with multiple predictors"
date: "due Tuesday, October 11 at 11:59p "
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
    toc: true
    toc_depth: 1
link-citations: yes
editor_options: 
  markdown: 
    wrap: sentence
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

In this lab we revisit the professor evaluations data we modelled in the previous AE.
Last time, we modeled evaluation scores using a single predictor at a time.
This time we will use multiple predictors to model evaluation scores.

For context, review the previous AE's introduction before continuing on to the exercises.

# Learning goals

-   Fitting a linear regression with multiple predictors
-   Interpreting regression output in context of the data
-   Comparing models

# Getting started

This lab will be completed collaboratively in your randomly assigned Labs 07 - 08 groups (see Canvas). The person randomly assigned to be the Group Leader in Canvas is responsible for submitting the assignment. 

+ Access the `MATH_250 Labs 07 - 08 G#` workspace via the link in the email invite you received. 
+ Launch the MATH_250 Labs 07 - 08` project
+ Open `lab-07.Rmd`

Remember that you are editing simultaneously and any changes you make in your files affect your whole group. Make sure to take turns so that each group member is "in the driver seat" for at least one exercise.

All plots should follow the best visualization practices discussed in lecture, including an informative title, labeled axes, and careful consideration ofaesthetic choices.

All code should follow the tidyverse style guidelines.

## Packages

We'll use the **tidyverse** package for much of the data wrangling and visualisation, the **tidymodels** package for modeling and inference, and the data lives in the **openintro** package.

```{r}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

## Data

The data can be found in the **openintro** package, and it's called `evals`.
Since the dataset is distributed with the package, we don't need to load it separately; it becomes available to us when we load the package.
You can find out more about the dataset by inspecting its documentation, which you can access by running `?evals` in the Console or using the Help menu in RStudio to search for `evals`.
You can also find this information [here](https://www.openintro.org/data/index.php?data=evals).

# Exercises

## Simple linear regression

1.  Fit a linear model (one you have fit before): `score_bty_fit`, predicting average professor evaluation `score` based on average beauty rating (`bty_avg`) only. Write the linear model, and note the $R^2$ and the adjusted $R^2$.

## Multiple linear regression

2.  Fit a new linear model: `score_bty_gen_fit`, predicting average professor evaluation `score` based on average beauty rating (`bty_avg`) and `gender`.
    Write the linear model, and note the $R^2$ and adjusted $R^2$.

3.  Interpret the slopes and intercept of `score_bty_gen_fit` in context of the data.

4.  What percent of the variability in `score` is explained by the model `score_bty_gen_fit`?

5.  What is the equation of the line corresponding to *just* male professors?

6.  For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

7.  How does the relationship between beauty and evaluation score vary between male and female professors?

8.  How do the adjusted $R^2$ values of `score_bty_gen_fit` and `score_bty_fit` compare?
    What does this tell us about how useful `gender` is in explaining the variability in evaluation scores when we already have information on the beauty score of the professor.

9. Compare the slopes of `bty_avg` under the two models (`score_bty_fit` and `score_bty_gen_fit`).
    Has the addition of `gender` to the model changed the parameter estimate (slope) for `bty_avg`?

10. Create a new model called `score_bty_rank_fit` with `gender` removed and `rank` added in.
    Write the equation of the linear model and interpret the slopes and intercept in context of the data.

# Grading (50pts)

| Component | Points |
|:----------|:-------|
| Ex 1      | 4      |
| Ex 2      | 4    |
| Ex 3      | 4   |
| Ex 4      | 4   |
| Ex 5      | 4      |
| Ex 6      | 4    |
| Ex 7      | 4   |
| Ex 8      | 4   |
| Ex 9      | 4   |
| Ex 10  | 4 |
| Workflow & formatting   | 10      |

**Grading notes**:

-   The "Workflow & formatting" grade is to assess the reproducible workflow. This includes having readable code (e.g. adequate use of spacing and line breaks), labeled code chunks, informative headers and sub-headers, and an overall organized and uncluttered report (e.g. suppress messages & warnings, no extraneous output).
