---
title: "Lab 03 - Nobel laureates"
subtitle: "due Tuesday, September 20 at 11:59p"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
    toc: true
    toc_depth: 1
link-citations: yes
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

In January 2017, Buzzfeed published an article on why Nobel laureates show immigration is so important for American science.
You can read the article [here](https://www.buzzfeednews.com/article/peteraldhous/immigration-and-science).
In the article they show that while most living Nobel laureates in the sciences are based in the US, many of them were born in other countries.
This is one reason why scientific leaders say that immigration is vital for progress.
In this lab we will work with the data from this article to recreate some of their visualizations as well as explore new questions.

# Learning goals

-   Replicating published results
-   Data wrangling and visualisation

# Lab prep

Prior to completing the lab, you should have read the Buzzfeed article titled [*These Nobel Prize Winners Show Why Immigration Is So Important For American Science*](https://www.buzzfeednews.com/article/peteraldhous/immigration-and-science)*.* We will replicate this analysis in the lab so it's crucial that you're familiar with it ahead of time.

# Getting started

This lab will be completed collaboratively in your randomly assigned Labs 03 - 04 groups (see Canvas). The person randomly assigned to be the Group Leader in Canvas is responsible for submitting the assignment. 

+ In Canvas, navigate to ["RStudio Cloud Links"](https://canvas.apu.edu/courses/42480/pages/rstudio-cloud-links) and click the link to join your group specific `MATH_250 Labs 03-04 G#` workspace. 
+ Launch the MATH_250 Labs 03 - 04` project
+ Open `lab-03.Rmd`

Remember that you are editing simultaneously and any changes you make in your files affect your whole group. Make sure to take turns so that each group member is "in the driver seat" for at least one exercise. 

## Warm up

Before we introduce the data, let's warm up with some simple exercises.

-   Update the YAML, changing the author name to your name, and **knit** the document.
-   Commit your changes with a meaningful commit message.
-   Push your changes to GitHub.
-   Go to your repo on GitHub and confirm that your changes are visible in your Rmd **and** md files. If anything is missing, commit and push again.

## Packages

We'll use the **tidyverse** package for much of the data wrangling.
This package is already installed for you.
You can load it by adding the following code to the `load-packages` code chunk in your `lab-03.Rmd` document:

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
```

## Data

The dataset for this assignment can be found as a CSV (comma separated values) file in the `data` folder of your repository.
You can read it in using the following.

```{r load-data, message = FALSE}
nobel <- read_csv("data/nobel.csv")
```

The variable descriptions are as follows:

-   `id`: ID number
-   `firstname`: First name of laureate
-   `surname`: Surname
-   `year`: Year prize won
-   `category`: Category of prize
-   `affiliation`: Affiliation of laureate
-   `city`: City of laureate in prize year
-   `country`: Country of laureate in prize year
-   `born_date`: Birth date of laureate
-   `died_date`: Death date of laureate
-   `gender`: Gender of laureate
-   `born_city`: City where laureate was born
-   `born_country`: Country where laureate was born
-   `born_country_code`: Code of country where laureate was born
-   `died_city`: City where laureate died
-   `died_country`: Country where laureate died
-   `died_country_code`: Code of country where laureate died
-   `overall_motivation`: Overall motivation for recognition
-   `share`: Number of other winners award is shared with
-   `motivation`: Motivation for recognition

In a few cases the name of the city/country changed after laureate was given (e.g. in 1975 Bosnia and Herzegovina was called the Socialist Federative Republic of Yugoslavia).
In these cases the variables below reflect a different name than their counterparts without the suffix \`\_original\`.

-   `born_country_original`: Original country where laureate was born
-   `born_city_original`: Original city where laureate was born
-   `died_country_original`: Original country where laureate died
-   `died_city_original`: Original city where laureate died
-   `city_original`: Original city where laureate lived at the time of winning the award
-   `country_original`: Original country where laureate lived at the time of winning the award

# Exercises

Take turns answering the exercises.
Make sure each team member has contributed by the time you submit your work.
You don't have to switch at each exercise, you can find your a cadence that works for your team and stick to it.

## Get to know your data

1.  How many observations and how many variables are in the dataset? Use inline code to answer this question. What does each row represent?

There are some observations in this dataset that we will exclude from our analysis to match the Buzzfeed results.

2.  Create a new data frame called `nobel_living` that filters for

-   laureates for whom `country` is available
-   laureates who are people as opposed to organizations (organizations are denoted with `"org"` as their `gender`)
-   laureates who are still alive (their `died_date` is `NA`)

```{r echo=FALSE}
nobel_living <- nobel %>%
  filter(
    !is.na(country),
    gender != "org",
    is.na(died_date)
  )
```

Confirm that once you have filtered for these characteristics you are left with a data frame with `r nrow(nobel_living)` observations, once again using inline code.

## Most living Nobel laureates were based in the US when they won their prizes

... says the Buzzfeed article.
Let's see if that's true.

First, we'll create a new variable to identify whether the laureate was in the US when they won their prize.
We'll use the `mutate()` function for this.
The following pipeline mutates the `nobel_living` data frame by adding a new variable called `country_us`.
We use an if statement to create this variable.
The first argument in the `if_else()` function we're using to write this if statement is the condition we're testing for.
If `country` is equal to `"USA"`, we set `country_us` to `"USA"`.
If not, we set the `country_us` to `"Other"`.

```{marginfigure}
Note that we can achieve the same result using the `fct_other()` function we've seen before (i.e. with `country_us = fct_other(country, "USA")`). We decided to use the `if_else()` here to show you one example of an if statement in R.
```

```{r}
nobel_living <- nobel_living %>%
  mutate(country_us = if_else(country == "USA", "USA", "Other"))
```

Next, we will limit our analysis to only the following categories: Physics, Medicine, Chemistry, and Economics.

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

For the next exercise work with the `nobel_living_science` data frame you created above.
This means you'll need to define this data frame in your R Markdown document, even though the next exercise doesn't explicitly ask you to do so.

3.  Create a faceted bar plot visualizing the relationship between the category of prize and whether the laureate was in the US when they won the nobel prize.
    Interpret your visualization, and say a few words about whether the Buzzfeed headline is supported by the data.

    -   Your visualization should be faceted by category.
    -   For each facet you should have two bars, one for winners in the US and one for Other.
    -   Flip the coordinates so the bars are horizontal, not vertical.

## But of those US-based Nobel laureates, many were born in other countries

```{marginfigure}
**Hint:** You should be able to ~~cheat~~ borrow from code you used earlier to create the `country_us` variable.
```

4.  Create a new variable called `born_country_us` that has the value `"USA"` if the laureate is born in the US, and `"Other"` otherwise.
    How many of the winners are born in the US?

5.  Add a second variable to your visualization from Exercise 3 based on whether the laureate was born in the US or not.
    Based on your visualization, do the data appear to support Buzzfeed's claim?
    Explain your reasoning in 1-2 sentences.

    -   Your final visualization should contain a facet for each category.
    -   Within each facet, there should be a bar for whether the laureate won the award in the US or not.
    -   Each bar should have segments for whether the laureate was born in the US or not.

## Here's where those immigrant Nobelists were born

```{marginfigure}
Note that your bar plot won't exactly match the one from the Buzzfeed article. This is likely because the data has been updated since the article was published.
```

6.  In a single pipeline, filter for laureates who won their prize in the US, but were born outside of the US, and then create a frequency table (with the `count()` function) for their birth country (`born_country`) and arrange the resulting data frame in descending order of number of observations for each country. Which country is the most common?

Now go back through your write up to make sure you've answered all questions and all of your R chunks are properly labelled.

# Submission

Once you decide as a team that you're done with this lab, to submit, you need to export your .html document from RStudio Cloud and then upload it to the Lab 03 assignment on Canvas. Refer back to [Lab 01](https://kgfitzgerald.github.io/APU_MATH_250/labs/lab-01-hello-r.html) for detailed instructions on how to export and submit your assignment. 

# Grading (50 pts)

<br>

| Component | Points |
|:----------|:-------|
| Ex 1      | 4      |
| Ex 2      | 8    |
| Ex 3      | 8    |
| Ex 4      | 6     |
| Ex 5      | 8      |
| Ex 6      | 8    |
| Workflow & formatting   | 8      |

**Grading notes**:

-   The "Workflow & formatting" grade is to assess the reproducible workflow. This includes having readable code (e.g. adequate use of spacing and line breaks), labeled code chunks, informative headers and sub-headers, and an overall organized and uncluttered report (e.g. suppress messages & warnings, no extraneous output). 

# Interested in how Buzzfeed made their visualizations?

The plots in the Buzzfeed article are called waffle plots.
You can find the code used for making these plots in Buzzfeed's GitHub repository [here](https://buzzfeednews.github.io/2017-01-immigration-and-science/).
You're not expected to recreate them as part of your assignment, but you're welcomed to do so for fun!
