---
title: "Lab 08 - The Office, Part II"
subtitle: "Feature Engineering & Cross Validation"
date: "due Tuesday, November 8 at 11:59pm"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
    toc: true
    toc_depth: 1
link-citations: yes
editor_options: 
  markdown: 
    wrap: sentence
---

```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Introduction

In today's lab you will analyze data from the [schrute](https://bradlindblad.github.io/schrute/) package to predict IMDB scores for episodes of The Office.

::: callout-note
This is the data that you began prepping at the end of AE-08. You may copy and paste your work into Exercises 1 - 4 below.
:::

### Learning goals

By the end of the lab you will...

-   engineer features based on episode scripts
-   train a model
-   interpret model coefficients
-   make predictions
-   evaluate model performance on training and testing data

## Getting started

This lab will be completed collaboratively in your randomly assigned Labs 07 - 08 groups (see Canvas). In RStudio Cloud, 

+ Navigate to the `MATH_250 Labs 07 - 08 G#` workspace 
+ Launch the MATH_250 Labs 07 - 08` project
+ Open `lab-08.qmd`

Remember that you are editing simultaneously and any changes you make in your files affect your whole group. Make sure to take turns so that each group member is "in the driver seat" for at least one exercise.

All plots should follow the best visualization practices discussed in lecture, including an informative title, labeled axes, and careful consideration ofaesthetic choices.

All code should follow the tidyverse style guidelines.

## Packages

The following packages are used in the lab.

```{r load-pacakges, message = FALSE}
library(tidyverse)
library(tidymodels)
library(schrute)
library(lubridate)
library(knitr)
```

## Data: The Office

The dataset for this lab comes from the [schrute](https://bradlindblad.github.io/schrute/) package and it's called `theoffice`.
This dataset contains the entire script transcriptions from The Office.

Let's start by taking a peek at the data.

```{r}
glimpse(theoffice)
```

```{r, echo = FALSE}
theoffice <- schrute::theoffice
```


There are 55130 observations and 12 columns in this dataset.
The variable names are as follows.

```{r}
names(theoffice)
```

Each row in the dataset is a line spoken by a character in a given episode of the show.
This means some information at the episode level (e.g., `imdb_rating`, `air_date`, etc. are repeated across the rows that belong to a single episode.

The `air_date` variable is coded as a factor, which is undesirable.
We'll want to parse that variable later into its components during feature engineering.
So, for now, let's convert it to date.

```{r}
theoffice <- theoffice %>%
  mutate(air_date = ymd(as.character(air_date)))
```

Let's take a look at the data to confirm we're happy with how each of the variables are encoded.

```{r}
glimpse(theoffice)
```

## Exercises

### Data prep

1. Identify episodes that touch on Halloween, Valentine's Day, and Christmas.

-   First, convert all `text` to lowercase with `str_to_lower()`.
-   Then, create three new variables (`halloween_mention`, `valentine_mention`, and `christmas_mention`) where that take on the value `1` if the character string `"halloween"`, `"valentine"`, or `"christmas"` appears in the text, respectively, and `0` otherwise.

Some code is provided below to help you get started.

```{r special-episodes, eval = FALSE}
theoffice <- theoffice %>%
  mutate(
    text = ___(text),
    halloween_mention = if_else(str_detect(text, "___"), ___, ___),
    valentine_mention = ___,
    ___ = ___
  )
```

2. In this exercise we'll accomplish two separate tasks.
And there's a good reason why we're doing it all at once; we're going to drastically change our data frame, from one row per line spoken to one row per episode.
We'll call the resulting data frame `office_episodes`.

The two tasks are as follows:

-   Task 1. Identify episodes where the word "halloween", "valentine", or "christmas" were ever mentioned, using variables you created above.
-   Task 2. Calculate the percentage of lines spoken by Jim, Pam, Michael, and Dwight for each episode of The Office.

Below are some instructions and starter code to get you started with these tasks.

-   Start by grouping `theoffice` data by `season`, `episode`, `episode_name`, `imdb_rating`, `total_votes`, and `air_date`. (These variables, except for `season` have the same value for each given episode, hence grouping by them allows us to make sure they appear in the output of this pipeline.)
-   Use `summarize()` to calculate the desired features at the season-episode level.
-   Task 1:
    -   Calculate the number of lines per season per episode, you might name this new variable `n_lines`.
    -   Then, calculate the proportion of lines in that episode spoken by each of the four `character`s Jim, Pam, Michael, and Dwight. Name these new variables `lines_jim`, `lines_pam`, `lines_michael`, and `lines_dwight`, respectively.
-   Task 2:
    -   Create a variable called `halloween` that sums up the `1`s in `halloween_mention` at the season-episode level and takes on the value `"yes"` if the sum is greater than or equal to 1, or `"no"` otherwise.
    -   Do something similar for new variables `valentine` and `christmas` as well based on values from `valentine_mention` and `christmas_mention`.
-   Finish up your `summarize()` statement by dropping the groups, so the resulting data frame is no longer grouped and remove `n_lines` (we won't use that variable in our analysis, we only calculated it as an intermediary step).

```{r episode-data, eval = FALSE}
office_episodes <- theoffice %>%
  group_by(___) %>%
  summarize(
    n_lines = n(),
    lines_jim = sum(character == "___") / n_lines,
    lines_pam = ___,
    lines_michael = ___,
    lines_dwight = ___,
    halloween = if_else(sum(___) >= 1, "yes", "no"),
    valentine = if_else(___, "___", "___"),
    christmas = if_else(___, "___", "___"),
    .groups = "drop"
  ) %>%
  select(-n_lines)
```

::: callout-note
Why `summarize()` and not `mutate()`?
We use `mutate()` to add / modify a column of a data frame.
The output data frame always has the same number of rows as the input data frame.
On the other hand, we use `summarize()` to reduce the data frame to either a single row (single summary statistic) or one row per each group (summary statistics at the group level).

And what about that `.groups` argument in summarize?
Try running your `summarize()` step without it first.
You'll see that R print out a message saying *"`summarize()` has grouped output by `season`, `episode`. You can override using the `.groups` argument."* `summarize()` will only drop the last group.
So if you want a data frame that doesn't have a grouping structure as a result of a `summaerize()`, you can explicitly ask for that with `.groups = "drop"`.
Before you proceed, read the [documentation for `summarize()`](https://dplyr.tidyverse.org/reference/summarise.html), and specifically the explanation for the `.groups` argument to prepare yourself for future instances where you might see this type of message.
:::

3. The Michael Scott character (played by Steve Carrell) left the show at the end of Season 7.
Add an indicator variable, `michael`, that takes on the value `"yes"` if Michael Scott (Steve Carrell) was in the show, and `"no"` if not.

```{r michael, eval = FALSE}
office_episodes <- office_episodes %>%
  mutate(michael = if_else(season > ___, "___", "___"))
```

4. Print out the dimensions (`dim()`) of the new dataset you created as well as the `names()` of the columns in the dataset.

Your new dataset, `office_episodes,` should have 186 rows and 14 columns.
The column names should be `season`, `episode`, `episode_name`, `imdb_rating`, `total_votes`, `air_date`, `lines_jim`, `lines_pam`, `lines_michael`, `lines_dwight`, `halloween`, `valentine`, `christmas`, and `michael`.
If you are not matching these numbers or columns, go back and try to figure out where you went wrong.
Or ask Dr. Fitz for help!

### Exploratory data analysis

This would be a good place to conduct some exploratory data analysis (EDA).
For example, plot the proportion of lines spoken by each character over time.
Or calculate the percentage of episodes that mention Halloween, or Valentine's Day, or Christmas. Given we have limited time in the lab we're not going to ask you to report EDA results as part of this lab, but we're noting this here to provide suggestions for how you might go about structuring your project.

### Modeling prep

5. Split the data into training (75%) and testing (25%).
Save the training and testing data as `office_train` and `office_test` respectively.

*Naming suggestion:* Call the initial split `office_split`, the training data `office_train`, and testing data `office_test`.

```{r split, eval = FALSE}
set.seed(123)
office_split <- ___(office_episodes)
office_train <- ___(office_split)
office_test <- ___(___)
```

6. Specify a linear regression model with engine `"lm"` and call it `office_spec`.

*Naming suggestion:* Call the model specification `office_spec`.

```{r model, eval = FALSE}
office_spec <- ___
```

7. Create a recipe that performs feature engineering using the following steps (in the given order):

-   `update_role()`: updates the role of `episode_name` to not be a predictor (be an ID)
-   `step_rm()`: removes `air_date` as a predictor
-   `step_dummy()`: creates dummy variables for `all_nominal_predictors()`
-   `step_zv()`: removes all zero variance predictors

*Naming suggestion:* Call the recipe `office_rec`.

```{r recipe, eval = FALSE}
office_rec <- recipe(imdb_rating ~ ., data = office_train) %>%
  ___
```

8. Build a model workflow for fitting the model specified earlier and using the recipe you developed to preprocess the data.

*Naming suggestion:* Call the model workflow `office_wflow`.

```{r wflow, eval = FALSE}
office_wflow <- workflow() %>%
  add_model(___) %>%
  add_recipe(___)
```

### Model fit and evaluation

9. Fit the model to training data, neatly display the model output, and interpret two of the slope coefficients.

*Naming suggestion:* Call the model fit `office_fit`.

```{r fit-tidy, eval = FALSE}
office_fit <- office_wflow %>%
  fit(data = ___)
___
```

10. Calculate predicted `imdb_rating` for the training data using the `predict()` function.
Then, bind two columns from the training data to this result: `imdb_rating` and `episode_name`.
The resulting data frame should have three columns: `.pred`, `imdb_rating`, and `episode_name`.
Then, using this data frame, create a scatterplot of predicted and observed IMDB ratings for the training data.

*Naming suggestion:* Call the resulting data frame `office_train_pred`.

*Stretch goal.* Add episode names, using `gghighlight()`, for episodes with much higher and much lower observed IMDB ratings compared to others.

11. Perform 5-fold cross-validation and view model performance metrics.


12. How does the model from this lab compare to the model fit in AE-08 in terms of performance? Which model do you think does a better job in predicting IMDB scores? 

13. Repeat Exercise 10, but with testing data.

*Naming suggestion:* Call the resulting data frame `office_test_pred`.

14. Report the R-squared and RMSE for this model for predictions on the testing data. Use these values to comment on the performance of this model. 

## Submission

Each group member should fill out this [group feedback form](https://qfreeaccountssjc1.az1.qualtrics.com/jfe/form/SV_1ZgcGyd7yfH095c) by the deadline.

Before submitting, open up your .html in a full window and inspect the following:

+ Check your code for neatness - add spaces and line breaks where appropriate to improve readability

+ Check any visualizations for informative and clean labels and titles

+ Check any tables for clean formatting (e.g. with `kbl() %>% kable_minmal()`)

+ Check that you've interpreted output as appropriate

+ Make sure extraneous warnings and messages are suppressed (e.g. set warning = FALSE, message = FALSE)

+ Check that Exercises are clearly labeled and your text responses are visually distinguished

+ Check for overall neatness & organization of the report (e.g. headers, subheaders, use of bullets, font changes)

To submit, you need to export your .html document from RStudio Cloud and then upload it to the Lab 08 assignment on Canvas. Refer back to [Lab 01](https://kgfitzgerald.github.io/APU_MATH_250/labs/lab-01-hello-r.html) for detailed instructions on how to export and submit your assignment. Recall only one group member needs to submit the assignment.

# Grading (50pts)

| Component | Points |
|:----------|:-------|
| Ex 1      | 3      |
| Ex 2      | 3    |
| Ex 3      | 3   |
| Ex 4      | 3   |
| Ex 5      | 3      |
| Ex 6      | 3    |
| Ex 7      | 3   |
| Ex 8      | 3   |
| Ex 9      | 3   |
| Ex 10  | 3 |
| Ex 11      | 3      |
| Ex 12      | 3    |
| Ex 13      | 3   |
| Ex 14      | 3   |
| Workflow & formatting   | 8      |

**Grading notes**:

-   The "Workflow & formatting" grade is to assess the reproducible workflow. This includes having readable code (e.g. adequate use of spacing and line breaks), labeled code chunks, informative headers and sub-headers, and an overall organized and uncluttered report (e.g. suppress messages & warnings, no extraneous output).
