---
title: "Lab 01: Hello R!"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: "tango"
    theme: "cosmo"
    df-print: paged
    fig-width: 3
    fig-height: 3
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(datasauRus)
```

## Reflection prompts

*Respond to each of the following after you have completed all exercises in this assignment*

*(RP1): What were the main concepts covered in this assignment?*

> YOUR ANSWER HERE

*(RP2): What's one thing you understand better after completing these exercises?*

> YOUR ANSWER HERE

*(RP3): What exercise(s) gave you the most trouble? What was difficult about them/where did you get stuck?*

> YOUR ANSWER HERE

# Exercise 1

```{r ex-1}

```

# Exercise 2

```{r ex-2}
```

# Exercise 3

```{r ex-3}
```

# Exercise 4

```{r ex-4}
```

# Exercise 5

```{r ex-5}

```
